Nix for Polyglot, Reproducible Data Science Workflows

### 2025-10-22 paper.qmd
Update paper.qmd

The package can then generate a visual representation of the pipeline's directed
acyclic graph by running `rxp_dag()`, as can be seen in Figure @fig-dag.

@@

I thank the rOpenSci reviewers and contributors who provided valuable feedback
on the development of [rix]{.pkg} and [rixpress]{.pkg}. In particular, I am
grateful to David Watkins and Jacob Wujiciak-Jens for their reviews of `{rix}`,
and to William Landau and Anthony Martinez for their reviews of `{rixpress}`. I
also acknowledge the contributions of Philipp Baumann, [rix]{.pkg}’s co-author,
Richard J. Acton, Jordi Rosell, Elio Campitelli, László Kupcsik, and 
Michael Heming for `{rix}`. Their expertise and feedback greatly 
improved the quality and usability of these packages.

@@

*   The exact versions of [R]{.proglang}, [Python]{.proglang}, and [Julia]{.proglang}.
*   A list of required packages for each language (e.g., `ggplot2` for R, `xgboost` for Python, `DataFrames` for Julia).
*   System-level dependencies like [Quarto]{.pkg}.

@@

*   **The RBC Model Simulation (`functions.jl`):** This [Julia]{.proglang} script contains a single pure
  function that implements the state-space solution to the RBC model, based on the economic theory of 
  @depaoli2009rbc. It takes the model's economic parameters as inputs and returns the simulated time-series
  data as a DataFrame.
*   **The XGBoost Forecasting Model (`functions.py`):** This [Python]{.proglang} script handles the 
  complete machine learning workflow through a series of modular functions. Its responsibilities include
  feature engineering (creating lagged variables), splitting the data into training and testing sets, 
  training the [XGBoost]{.pkg} model, and generating predictions.
*   **The Visualization (`functions.R`):** This [R]{.proglang} script contains a function that uses
  `[ggplot2]{.pkg}` to create the final visualization. It is designed to take the data frame of 
  actual and predicted values produced by the Python script and generate a plot comparing the two series.

@@

The entire workflow is defined and orchestrated by `gen-pipeline.R`. This script acts as the master plan, 
using functions from the `[rixpress]{.pkg}` package to define each computational step as a *derivation*.
It declaratively outlines the dependencies between steps:

1.  It begins by defining the RBC model parameters in [Julia]{.proglang}.
2.  It specifies that the **RBC simulation** in `functions.jl` depends on these parameters.
3.  It then defines the series of **Python steps** (feature preparation, training, prediction), 
  making each one dependent on the output of the previous one. The first Python step is explicitly 
  made dependent on the output from the Julia simulation. `rixpress` handles the passing of data 
  between languages, in this case using the Arrow file format for efficiency.
4.  Finally, it defines the **R visualization** step, which depends on the final predictions from 
  the Python model, and a **Quarto rendering** step that depends on the generated plot.

>>>>

The package can then generate a visual representation of the pipeline's directed
acyclic graph by running `rxp_dag()`, as can be seen in @fig-dag.

@@

I thank the rOpenSci reviewers and contributors who provided valuable feedback
on the development of [rix]{.pkg} and [rixpress]{.pkg}. In particular, I am
grateful to David Watkins and Jacob Wujiciak-Jens for their reviews of [rix]{.pkg},
and to William Landau and Anthony Martinez for their reviews of [rixpress]{.pkg}. I
also acknowledge the contributions of Philipp Baumann, [rix]{.pkg}’s co-author,
Richard J. Acton, Jordi Rosell, Elio Campitelli, László Kupcsik, and 
Michael Heming for [rix]{.pkg}. Their expertise and feedback greatly 
improved the quality and usability of these packages.

@@

*   The exact versions of [R]{.proglang}, [Python]{.proglang}, and [Julia]{.proglang}.
*   A list of required packages for each language 
  (e.g., [ggplot2]{.pkg} for [R]{.proglang}, [xgboost]{.pkg} for [Python]{.proglang}, [DataFrames]{.pkg} for [Julia]{.proglang}).
*   System-level dependencies like [Quarto]{.pkg}.

@@

*   The RBC Model Simulation (`functions.jl`): This [Julia]{.proglang} script contains a single pure
  function that implements the state-space solution to the RBC model, based on @depaoli2009rbc.
  It takes the model's economic parameters as inputs and returns the simulated time-series
  data as a DataFrame.
*   The XGBoost Forecasting Model (`functions.py`): This [Python]{.proglang} script handles the 
  complete machine learning workflow through a series of modular functions. Its responsibilities include
  feature engineering (creating lagged variables), splitting the data into training and testing sets, 
  training the [XGBoost]{.pkg} model, and generating predictions.
*   The Visualization (`functions.R`): This [R]{.proglang} script contains a function that uses
  [ggplot2]{.pkg} to create the final visualization. It is designed to take the data frame of 
  actual and predicted values produced by the Python script and generate a plot comparing the two series.

@@

The entire workflow is defined and orchestrated by `gen-pipeline.R`. This script acts as the master plan, 
using functions from the [rixpress]{.pkg} package to define each computational step as a *derivation*.
It declaratively outlines the dependencies between steps:

1.  It begins by defining the RBC model parameters in [Julia]{.proglang}.
2.  It specifies that the **RBC simulation** in `functions.jl` depends on these parameters.
3.  It then defines the series of **Python steps** (feature preparation, training, prediction), 
  making each one dependent on the output of the previous one. The first Python step is explicitly 
  made dependent on the output from the Julia simulation. [rixpress]{.pkg} handles the passing of data 
  between languages, in this case using the Arrow file format for efficiency.
4.  Finally, it defines the **R visualization** step, which depends on the final predictions from 
  the Python model, and a **Quarto rendering** step that depends on the generated plot.
